ms_2_results

### 2018-01-10 man/ms_1_material_methods.Rmd
update disturbances

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and mean, minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010, @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We select a temporal scale of 6 months.

@@

>>>>

una frase conectora 
We also explored the climate-growth relationships and the disturbance  ...

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and mean, minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. Data were downloaded and preprocess using the climate4R bundle (http://www.meteo.unican.es/en/climate4R) [@Cofino2018, @Frias2018].
We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010, @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We select a temporal scale of 6 months.

@@

### Disturbance analyses

Disturbance chronologies were built using tree-ring width to identify abrupt and sustained increases (releases) or decreases (supressions) in radial growth [@Nowacki1997]. ... This approach minimizes the long-term growth trends and interannual growth variations usually driven by climate, while enhances decadal abrupt and sustained radial-growth changes characteristic of forest disturbances.
By using a temporal window of reasonable length, this method filters out the response to short-term changes in temperature and precipitation [@Nowacki1997, @Fraver2005].

Growth changes (GC) were calculated for the individual tree-ring series using a 10-year running window as either positive (PGC) or negative (NGC) growth changes:

$$\mathrm{\% GC} = \left [ \frac{\left ( M1 - M2 \right )}{M2}\right] \times 100 $$
where $M1$ is preceding 10-year median and $M2$ is subsequent 10-year median. We used medians since they are mores robust estimator of central tendency than means @Rubino2004 @Camarero2011]. Site disturbance chronologies were constructed by averaging the individual disturbances series annually. A

To separate growth periods produced by disturbance events from those by climate, we considered a treshold of 25 % of GC for consi

A minimum $GC$ of 25% was established to depict canopy disturbances. Furthermore, more than 50% of the individual trees displaying the same growth changes was considered a stand-wise disturbance.

This approach filters out

We used the percentage Growth-change filter (GC).

, calculated annually using 10-year windows

Disturbance chronologies were built using tree-ring width to identify historical abrupt positive (releases) or negative (suppressions) changes in growth in each species (Nowacki and Abrams, 1997). This approach minimizes the long-term growth trends and interannual growth variations usually driven by climate, while enhances decadal abrupt and sustained radial-growth changes characteristic of forest disturbances.

### 2018-01-10 man/ms_2_results.Rmd
update
update disturbances

```{r setup, echo=FALSE}
show_text <- FALSE
```

@@

>>>>

```{r, echo=FALSE}
show_text <- FALSE
```

@@

# Climate and tree

Tree-growth exhibited a signigicant response to water availability more than to the temperatures. Cumulative precipitation of the hydrological year (September of the previous year to August of the growth year) influenced positively the tree-growth for the both northern and southern populations (FIG ). Precipitation of previous december was also positively correlated with tree growth in the northern population and in the highest location of the southern population. Hydrological, Spring and Summer SPEI showed a strong positive correlation with tree-growth, specially for northern population. Spring Maximum temperature was the most significant limitation of the tree-growth for the southern population while maximum temperatures of the previous August did for the northern one. Conversely, mean, minimum and maximum temperatures of current September positively influenced the tree-growth in the northern population.
